---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# plFA

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/plFA)](https://CRAN.R-project.org/package=plFA)
[![R-CMD-check](https://github.com/giuseppealfonzetti/plFA/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/giuseppealfonzetti/plFA/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/giuseppealfonzetti/plFA/graph/badge.svg)](https://app.codecov.io/gh/giuseppealfonzetti/plFA)
<!-- badges: end -->

The `{plFA}` package allows the estimation of confirmatory factor models for ordinal data using stochastic and numeric pairwise likelihood optimisation. 

## Installation

You can install the development version of plFA from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("giuseppealfonzetti/plFA")
```

When the package is installed, you can load it with:

```{r}
library(plFA)
```

The startup message indicates that the `{plFA}` package provides its own version of the `cfa()` function, which can be used to estimate plFA models using the `lavaan` syntax.
An example is given below.

## Fitting an ordinal CFA model

Consider the data containing 25 personality items measured on 6-point scale from the International Personality Item Pool (IPIP) dataset.
The data set is available in the `{plFA}` package and can be loaded with the following code:

```{r}
data("bfi")
head(bfi)
```

The `{lavaan}` model syntax to fit a confirmatory factor analysis model with 5 latent factors, where the first 5 items load on the first factor and so on, is:

```{r}
mod <- "
  opn =~ O1 + O2 + O3 + O4 + O5  # Openness
  con =~ C1 + C2 + C3 + C4 + C5  # Conscientiousness
  ext =~ E1 + E2 + E3 + E4 + E5  # Extraversion
  agr =~ A1 + A2 + A3 + A4 + A5  # Agreeableness
  neu =~ N1 + N2 + N3 + N4 + N5  # Neuroticism
"
```

Users of `{lavaan}` will be familiar with this syntax. 
The call to the `cfa()` function is made as usual:

```{r}
# Fit the model (note, this cfa() function is from the plFA package)
fit <- cfa(model = mod, data = bfi, std.lv = TRUE, test = "mean.var.adjusted")
summary(fit)
```


All^[Most :)--we're working on it!] `lavaan` methods should carry over, such as:

```{r}
head(coef(fit), 10)
head(parameterestimates(fit, se = TRUE))
head(fitted(fit)$cov)  # model-implied covariance matrix for the UVs
```

And also plot from the `{semPlot}` package:

```{r}
#| label: fig-bfi
#| out-width: 100%
#| dpi: 300
library(semPlot)

grouplist <- list(
  lat = c("opn", "con", "ext", "agr", "neu"),
  man = rep(c("opn", "con", "ext", "agr", "neu"), each = 5)
)

# Define a named vector of colors for each group (latent factor)
groupcols <- RColorBrewer::brewer.pal(5, "Set2") |>
  rep(each = 5) |>
  adjustcolor(alpha.f = 0.5)

# Now call semPaths
semPaths(fit, whatLabels = "est", layout = "circle", residuals = FALSE, 
         intercept = FALSE, groups = grouplist, color = groupcols, 
         legend = FALSE)
```

## Session information

```{r}
#| label: sessioninfo
sessioninfo::session_info()
```

<!-- To fit a latent factor analysis model with 5 factors, where the first 5 items load on the first factor and so on, we can use the `fit_plFA()` function. -->
<!-- In preparation, we must build the following objects to be supplied as a list to the `CONSTR_LIST` argument: -->

<!-- - `CONSTRMAT`: A matrix of constraints on the factor loadings. -->
<!-- - `CORRFLAG`: A flag indicating whether the factor correlations are estimated. -->
<!-- - `CONSTRVAR`: A vector of constraints on the factor variances, if any. -->
<!-- - `STDLV`: A flag indicating whether the factor loadings are standardised. -->
<!-- - `LLC`: A list indicating linear constraints on factor loading estimations, if any. -->

<!-- The above ingredients are prepared using the code below. -->

<!-- ```{r} -->
<!-- p <- 25         # number of items -->
<!-- q <- 5          # number of factors -->
<!-- n <- nrow(bfi)  # number of observations -->

<!-- # Convert data frame to matrix, with each item taking values 0,1,2,...,5 -->
<!-- D <- as.matrix(data.frame(lapply(bfi[, 1:p], as.numeric))) - 1 -->
<!-- head(D) -->

<!-- # Constraint matrix: NA indicate free loadings -->
<!-- constrmat <- build_constrMat(P = p, Q = q, STRUCT = "simple") -->
<!-- head(constrmat, 12) -->

<!-- # Build constraint list -->
<!-- constr_list <- list( -->
<!--   CONSTRMAT = constrmat, -->
<!--   CORRFLAG = TRUE, -->
<!--   CONSTRVAR = rep(1, q),  # these options standardise the   -->
<!--   STDLV = TRUE,           # factor variances -->
<!--   LLC = NULL -->
<!-- ) -->
<!-- ``` -->

<!-- We can now fit the factor model: -->

<!-- ```{r} -->
<!-- fit <- fit_plFA( -->
<!--   DATA = D,  -->
<!--   CONSTR_LIST = constr_list, -->
<!--   METHOD = "ucminf", -->
<!--   VERBOSE = TRUE -->
<!-- ) -->
<!-- print(fit) -->
<!-- ``` -->

<!-- The default method is using the optimiser from the `{ucminf}` package, which works well for small data sizes. -->
<!-- Alternatively, the user may choose `METHOD = "SA"` to turn on the stochastic approximation algorithm. -->
<!-- Use the function `getPar()` to extract the parameter estimates: -->

<!-- ```{r} -->
<!-- getPar(fit) -->
<!-- ``` -->

<!-- Computing the standard errors is done separately using the `computeVar()` function: -->

<!-- ```{r} -->
<!-- var <- computeVar(OBJ = fit, DATA = D, VERBOSE = TRUE) -->
<!-- str(var) -->

<!-- # Standard errors -->
<!-- head(sqrt(diag(var$vcov) / n)) -->
<!-- ``` -->

<!-- ## `{lavaan}` wrapper -->

<!-- This package also provides a user-friendly `{lavaan}`-style interface to estimate factor analysis models using the `cfa()` function. -->
<!-- Let's rerun the above example, but this time using the stochastic approximation algorithm. -->
