---
title: "Limited information Goodness-of-Fit tests"
author: Haziq Jamil
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Limited information Goodness-of-Fit tests}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(tidyverse)
library(tinytest)
```

## Limited information GOF tests

The `{lavaan.bingof}` package provides a set of LIGOF tests for CFA models with binary data.
Models estimated using the `{plFA}` package are directly compatible.

```{r setup}
library(plFA)
library(lavaan.bingof)
```

```{r}
(dat <- gen_data_bin(n = 1000, seed = 123))
mod <- "eta =~ y1 + y2 + y3 + y4 + y5"
fit <- cfa(mod, dat, std.lv = TRUE)
expect_true(inherits(fit, "plFAlavaan"))

# Test statistics
all_tests(fit)
```

As a comparison, we can look at the output from using the `cfa()` function in the `{lavaan}` package.

```{r}
fit1 <- plFA::cfa(mod, dat, std.lv = TRUE)
fit2 <- lavaan::cfa(mod, dat, std.lv = TRUE, estimator = "PML")

# Test equality of coefficients and standard errors
expect_equal(coef(fit1), coef(fit2), tolerance = 1e-5)  
expect_equal(fit1@Fit@se, fit2@Fit@se, tolerance = 1e-5)

# Test equality of test statistics and p-values
c(
  "Wald_test",
  "Wald_vcovf_test",
  "Wald_diag_test",
  "Pearson_test",
  "RSS_test",
  "Multn_test"
) |>
  set_names() |>
  map(.f = function(x) {
    out1 <- do.call(x, list(fit1))
    out2 <- do.call(x, list(fit2))
    expect_equal(out1, out2, tolerance = 5e-3)
  })
```

## Testing differences


```{r}
information_matrix <- function(fit, type) {
  lavargs <- list(
    lavmodel = fit@Model,
    lavsamplestats = fit@SampleStats,
    lavdata = fit@Data,
    lavoptions = fit@Options,
    lavcache = fit@Cache,
    inverted = FALSE
  )

  if (type == "observed") 
    out <- do.call("lav_model_information_observed", lavargs, 
                   envir = asNamespace("lavaan"))
  else if (type == "expected")
    out <- do.call("lav_model_information_expected", lavargs, 
                   envir = asNamespace("lavaan"))
  else if (type == "first.order")
    out <- do.call("lav_model_information_firstorder", lavargs, 
                   envir = asNamespace("lavaan"))
  else stop("Invalid type")
  return(out)
}
```


### Sensitivity (inverted) matrix


```{r}
# From plFA
Hinv1 <- with(fit1@external, computeVar(plFA, D)$invH[idx_plFA2lav, idx_plFA2lav])
fit1 <- cfa(mod, dat, std.lv = TRUE, estimator.args = list(computevar_numderiv = TRUE), verbose = TRUE)
Hinv2 <- with(fit1@external, computeVar(plFA, D)$invH[idx_plFA2lav, idx_plFA2lav])
Hinv3 <- solve(information_matrix(fit1, "observed"))

expect_equal(Hinv1, Hinv2, tolerance = 1e-5)  # ucminf vs numderiv
expect_equal(Hinv1, Hinv3, tolerance = 1e-5)  # ucminf vs lavaan
expect_equal(Hinv2, Hinv3, tolerance = 1e-5)  # numderiv vs lavaan
```

### Variability matrix

```{r}
# From plFA
J1 <- with(fit1@external, computeVar(plFA, D)$J[idx_plFA2lav, idx_plFA2lav])
J2 <- information_matrix(fit1, "first.order")

expect_equal(J1, J2, tolerance = 1e-5)  # ucminf vs numderiv
```

## Session information

```{r}
sessioninfo::session_info()
```
